---
title: "Run_analysis"
author: "Michael Volk"
date: "May 3, 2020"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages, include=FALSE}
library(dplyr)
```

## Import Data
```{r}
activity_labels <- read.table("C:/Users/michaelvolk/Desktop/datasciencecoursera/R/Getting_and_Cleaning_Data/UCI HAR Dataset/activity_labels.txt")
features <- read.table("C:/Users/michaelvolk/Desktop/datasciencecoursera/R/Getting_and_Cleaning_Data/UCI HAR Dataset/features.txt")
train_X <- read.table("C:/Users/michaelvolk/Desktop/datasciencecoursera/R/Getting_and_Cleaning_Data/UCI HAR Dataset/train/X_train.txt")
train_y <- read.table("C:/Users/michaelvolk/Desktop/datasciencecoursera/R/Getting_and_Cleaning_Data/UCI HAR Dataset/train/y_train.txt")
train_subject <- read.table("C:/Users/michaelvolk/Desktop/datasciencecoursera/R/Getting_and_Cleaning_Data/UCI HAR Dataset/train/subject_train.txt")
test_X <- read.table("C:/Users/michaelvolk/Desktop/datasciencecoursera/R/Getting_and_Cleaning_Data/UCI HAR Dataset/test/X_test.txt")
test_y <- read.table("C:/Users/michaelvolk/Desktop/datasciencecoursera/R/Getting_and_Cleaning_Data/UCI HAR Dataset/test/y_test.txt")
test_subject <- read.table("C:/Users/michaelvolk/Desktop/datasciencecoursera/R/Getting_and_Cleaning_Data/UCI HAR Dataset/test/subject_test.txt")
```

## Define variable for subsetting
*column_subset* is used to find all of the column variables that represent a previously computed mean or standard deviation. Importatnly the "-" character helps to specify to occurences where the variable is **Not** a function of the mean or standard deviation. 
```{r}
#Keep only columns with mean and std, without selecting f(mean|std)
column_subset <- grep("-[Mm]ean|-[Ss]td", features[[2]])
```

## Combine subjects, data, and labels for test and training
Here we append the unique subject identifying vector to the left hand side of the data tables, we rename the column headings so they match the features.txt, we add labels to the dataset, and we take the subset of the data to capture vectors representing means and standard deviations.
```{r}
#Treat train data
colnames(train_X) <- features[[2]]
colnames(train_y)<- "label_num"
train_X <- train_X[,column_subset]
train <- cbind(train_X, train_y)
train <- cbind(train_subject, train)
```
We repeat the same procedure here for the test data.
```{r}
#Treat test data
colnames(test_X) <- features[[2]]
colnames(test_y)<- "label_num"
test_X <- test_X[,column_subset]
test <- cbind(test_X, test_y)
test <- cbind(test_subject, test)
```
We now combine the test and train data
```{r}
#Create combined data table
data <- rbind(train, test)
```

## First tidy datatset
The subject column is propely labled as it was the only column without a proper label. Now we add a vector for activity labels that corresponds with the *label_num* variable. Then we drop the *label_num* variable as to not be redundant.
```{r}
colnames(data)[1] <- "subject"
data_tidy1 <- data %>% 
      mutate(label = activity_labels[[2]][label_num]) %>% 
      select(-(label_num))
print(tbl_df(data_tidy1))
```

## Second tidy dataset
First we group the dataset by subject and label so the mean will be taken across variables that share group headers. Next we collapse the table to display the means with the summarize function, and then we arrange by subject.
```{r}
data_tidy2 <- data_tidy1 %>%
      group_by(subject, label) %>% 
      summarize_all(mean) %>% 
      arrange(subject)
print(data_tidy2)
```