---
title: "Final_Project needs a real title"
author: "Michael Volk"
date: "May 26, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Synopsis

_describe and summarize the analysis in at most 10 complete sentences_ 

## Reading Documentation

### Questions 1:

Across the United States, which types of events are most harmful with respect to population health?

Initial thoughts for exploration:

- types of storms and direct fatality

### Question 2:

Across the United States, which types of events have the greatest economic consequences?

## Data Processing

Read in data

```{r read_in_data, cache = TRUE}
df <- read.csv("repdata_data_StormData.csv.bz2")
```

Table processing for figures

```{r table_1, cache = TRUE}
fatalities_months <- df %>% mutate(MONTHS = fct_explicit_na(months(as.Date(df$BGN_DATE, "%m/%d/%y")))) %>%
      group_by(EVTYPE, MONTHS) %>%
      summarize(FATALITIES_sum = sum(FATALITIES)) %>%
      arrange(desc(FATALITIES_sum)) %>% 
      ungroup()
```
## Results

```{r figure_1}
n_obs <- 100
fatalities_months_short <- fatalities_months[1:n_obs,]
#reordering factors - Months: chronologically, Event types: Months with highest fatality totals
fatalities_months_short$MONTHS <- factor(fatalities_months_short$MONTHS, levels = month.name)
fatalities_months_short$EVTYPE <- factor(fatalities_months_short$EVTYPE, levels = unique(fatalities_months_short$EVTYPE))

g1 <- ggplot(fatalities_months_short, aes(EVTYPE, FATALITIES_sum)) + 
      geom_point(aes(color = MONTHS), size = 4, alpha = 0.3) + 
      theme(axis.text.x = element_text(angle = 45, hjust=1), plot.title = element_text(hjust = 0.5)) +
      scale_y_continuous(trans = "log10") +
      labs(title = "Storm Event Fatalities by Month", x = "Event Type", y = expression(Total~Fatalities~(log[10]~scale)))
```



```{r}


```








